Quantum Hoare Type Theory

### 2020-10-09 ms-paper/ms-paper.tex
Move previous derivation to appendix
Use correct notation for pair type

@@

\lstset{
	basicstyle=\small\ttfamily,
	flexiblecolumns=false,
	basewidth={0.5em,0.45em},
	commentstyle=\color{dkgreen},
	keywordstyle=\bfseries\color{blue},
	stringstyle=\color{violet},
	literate={+}{{$+$}}1 {/}{{$/$}}1 {*}{{$*$}}1 {=}{{$=$}}1
	{>}{{$>$}}1 {<}{{$<$}}1 {\\}{{$\lambda$}}1
	{\\\\}{{\char`\\\char`\\}}1
	{->}{{$\rightarrow$}}2 {>=}{{$\geq$}}2 {<-}{{$\leftarrow$}}2
	{<=}{{$\Leftarrow$}}2 {=>}{{$\Rightarrow$}}2
	{\ .}{{$\circ$}}2 {\ .\ }{{$\circ$}}2
	{>>}{{>>}}2 {>>=}{{>>=}}2
	{|}{{$\vert$}}1 {\\>}{{$\rangle$}}1
	{/\\}{{$\wedge$}}1 {\\/}{{$\vee$}}1
	{\\in}{{$\in$}}1 {\\o}{{$\circ$}}1 {\\Pi}{{$\Pi$}}1
	{\\Phi+}{{$\Phi^+$}}2
	{|->}{{$\mapsto$}}2 {-o}{{$\multimap$}}2
}

@@

WIP derivation tree for above program:

\begin{mathpar}
	\inferrule*[right=HIntro]
	{\inferrule*[right=consq]
		{\inferrule*[right=mkQbit]
			{\inferrule*[right=measQbit]
				{\inferrule*[right=return]
					{q : \type{Qbit}, r : \type{Bool} \vdash r \chk \type{Bool}}
					{q : \type{Qbit}, r : \type{Bool}; P_1 \circ ((q \mapsto -) \multimap emp) \vdash \kw{return} r \syn r : \type{Bool} . R}
					\\ q : \type{Qbit} \vdash q \chk \type{Qbit}
					\\ \cdot; P_0 * (q \mapsto \ket{0}) \prove (q \hookrightarrow -)
				}
				{q : \type{Qbit}; P_0 * (q \mapsto \ket{0}) \vdash r \Leftarrow \kw{measQbit} q; E_3 \syn r : \type{Bool} . (\exists r : \type{Bool}.R)}
				\\ \cdot \vdash \kw{false} \chk \type{Bool}
			}
			{
				\cdot; P_0 \vdash q \Leftarrow \kw{mkQbit false}; E_{23} \syn r : \type{Bool} . (\exists q : \type{Qbit}.R)
			}
			\\ r : \type{Bool}; i,m ; R \prove emp \multimap emp \wedge Id(r, \kw{false})
		}
		{
			\cdot; \lambda i . \lambda m . i = m \wedge (emp * \lambda m' \top)m \vdash E_1; E_{23} \chk r : \type{Bool} . (emp \multimap emp \wedge Id(r, \kw{false}))
		}
	}
	{\cdot \vdash \kw{do} E \chk \{emp\} r : \type{Bool} \{emp \wedge \kw{Id}(r, \kw{false})\}}
\end{mathpar}

@@

@@

\subsection{Bell pair}
\todo{TODO: The type should be \type{Qbit} $\otimes$ \type{Qbit}}
\begin{lstlisting}[language=QHaskell]
bell : {emp} (a, b) : (Qbit, Qbit) {Id(a, b) /\ a \in {|0\>, |1\>}}
     = do qa <- qplus;
          qb <- share qa;
          return (qa, qb)
\end{lstlisting}

\subsection{Testing modular Bell pair}
\begin{lstlisting}[language=QHaskell]
testBell : {emp} (a, b) : (Bool, Bool) {emp /\ Id(a, b)}
         = do (qa, qb) <- bell;
              return (measQbit qa, measQbit qb)
\end{lstlisting}

\section{Teleportation}
\label{app:teleport}
\todo{TODO: Refine assertions}
\subsection{Alice's circuit}
\begin{lstlisting}[language=QHaskell]
alice : \Pi a : Qbit. \Pi e : Qbit
        {(Id(a, -) /\ entangled(e)}
           r : (Bool, Bool)
        {emp}
      = \a.\e.do applyU (ifQ a (X e));
                 applyU (H a);
                 return (measQbit a, measQbit e)
\end{lstlisting}

>>>>

\usepackage{rotating}

@@

\lstset{
	basicstyle=\small\ttfamily,
	flexiblecolumns=false,
	basewidth={0.5em,0.45em},
	commentstyle=\color{dkgreen},
	keywordstyle=\bfseries\color{blue},
	stringstyle=\color{violet},
	literate={+}{{$+$}}1 {/}{{$/$}}1 {*}{{$*$}}1 {=}{{$=$}}1
	{>}{{$>$}}1 {<}{{$<$}}1 {\\}{{$\lambda$}}1
	{\\\\}{{\char`\\\char`\\}}1
	{->}{{$\rightarrow$}}2 {>=}{{$\geq$}}2 {<-}{{$\leftarrow$}}2
	{<=}{{$\Leftarrow$}}2 {=>}{{$\Rightarrow$}}2
	{\ .}{{$\circ$}}2 {\ .\ }{{$\circ$}}2
	{>>}{{>>}}2 {>>=}{{>>=}}2
	{|}{{$\vert$}}1 {\\>}{{$\rangle$}}1
	{/\\}{{$\wedge$}}1 {\\/}{{$\vee$}}1
	{\\in}{{$\in$}}1 {\\o}{{$\circ$}}1 {\\Pi}{{$\Pi$}}1
	{\\Phi+}{{$\Phi^+$}}2
	{|->}{{$\mapsto$}}2 {-o}{{$\multimap$}}2
	{o*}{{$\otimes$}}1
}

@@

@@

\section{Sample Proof}
Complete derivation tree for Hello Quantum World:

\begin{turn}{90}
	\begin{minipage}{40pc}
	\begin{mathpar}
		\inferrule*[right=HIntro]
		{\inferrule*[right=consq]
			{\inferrule*[right=mkQbit]
				{\inferrule*[right=measQbit]
					{\inferrule*[right=return]
						{q : \type{Qbit}, r : \type{Bool} \vdash r \chk \type{Bool}}
						{q : \type{Qbit}, r : \type{Bool}; P_1 \circ ((q \mapsto -) \multimap emp) \vdash \kw{return} r \syn r : \type{Bool} . R}
						\\ q : \type{Qbit} \vdash q \chk \type{Qbit}
						\\ \cdot; P_0 * (q \mapsto \ket{0}) \prove (q \hookrightarrow -)
					}
					{q : \type{Qbit}; P_0 * (q \mapsto \ket{0}) \vdash r \Leftarrow \kw{measQbit} q; E_3 \syn r : \type{Bool} . (\exists r : \type{Bool}.R)}
					\\ \cdot \vdash \kw{false} \chk \type{Bool}
				}
				{
					\cdot; P_0 \vdash q \Leftarrow \kw{mkQbit false}; E_{23} \syn r : \type{Bool} . (\exists q : \type{Qbit}.R)
				}
				\\ r : \type{Bool}; i,m ; R \prove emp \multimap emp \wedge Id(r, \kw{false})
			}
			{
				\cdot; \lambda i . \lambda m . i = m \wedge (emp * \lambda m' \top)m \vdash E_1; E_{23} \chk r : \type{Bool} . (emp \multimap emp \wedge Id(r, \kw{false}))
			}
		}
		{\cdot \vdash \kw{do} E \chk \{emp\} r : \type{Bool} \{emp \wedge \kw{Id}(r, \kw{false})\}}
	\end{mathpar}
	\end{minipage}
\end{turn}

@@

\subsection{Bell pair}
\begin{lstlisting}[language=QHaskell]
bell : {emp} (a, b) : (Qbit o* Qbit) {Id(a, b) /\ a \in {|0\>, |1\>}}
     = do qa <- qplus;
          qb <- share qa;
          return (qa, qb)
\end{lstlisting}

\subsection{Testing modular Bell pair}
\begin{lstlisting}[language=QHaskell]
testBell : {emp} (a, b) : (Bool o* Bool) {emp /\ Id(a, b)}
         = do (qa, qb) <- bell;
              return (measQbit qa, measQbit qb)
\end{lstlisting}

\section{Teleportation}
\label{app:teleport}
\todo{TODO: Refine assertions}
\subsection{Alice's circuit}
\begin{lstlisting}[language=QHaskell]
alice : \Pi a : Qbit. \Pi e : Qbit
        {(Id(a, -) /\ entangled(e)}
           r : (Bool o* Bool)
        {emp}
      = \a.\e.do applyU (ifQ a (X e));
                 applyU (H a);
                 return (measQbit a, measQbit e)
\end{lstlisting}
